---
title: "Plotting HOBO outputs"
output: html_document
date: "2024-01-08"
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(data.table)
library(lubridate)
```

Load data
```{r}
# Example data
# from .hobo file that was exported to .csv in the HOBOware program
data <- fread("10596690_Birds_Mammals_Range.csv", skip=1, header=T)
data <- fread("20240109_Birds_Mammals_fluid_rm.csv", skip=1, header=T)
data <- fread("Temp_Birds_Mammals_fluid_rm_20240207.csv", skip=1, header=T)


colnames(data)  <- c("n","timestamp", "temp", "rh", "dewpt", "host", "end")

# change date/time to POSIXct class
data$time <- as.POSIXct(data$timestamp, format = "%m/%d/%y  %I:%M:%S %p")

# get just a date class for future filtering
data$date <- as.Date(data$time, format = "%m/%d/%Y")


min(data$date) 
max(data$date)
```


Plot
1) Temperature

* Add temp min 65 max 75 lines (hline) - done
* add function that highlights time periods where the value exceeds that threshold (rect)

```{r}
# Temperature
ggplot(data, aes(x=date, y=temp)) +
  geom_line() +
  geom_smooth(col="red") +
  geom_hline(aes(yintercept = 62), color="red", linetype="dashed") +
  geom_hline(aes(yintercept = 70), color="red", linetype="dashed") +
  labs(x="Date", y="Temperature") +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B")
```

### Zoom in
```{r}
# Temperature
ggplot(data, aes(x=date, y=temp)) +
  geom_line() +
  geom_smooth(col="red") +
  geom_hline(aes(yintercept = 62), color="red", linetype="dashed") +
  geom_hline(aes(yintercept = 70), color="red", linetype="dashed") +
  labs(x="Date", y="Temperature") +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B",
               limits = as.Date(c("2023-12-14", "2024-02-08")))
```



2) Relative humidity
```{r}
ggplot(data, aes(x=time, y=rh)) +
  geom_line() +
  geom_smooth(col="blue") +
  geom_hline(aes(yintercept = 30), color="blue", linetype="dashed") +
  geom_hline(aes(yintercept = 50), color="blue", linetype="dashed") +
  labs(x="Date", y="Temperature")
```


3) Plot all 3 (Temp, RH, Dew pt)

ggplot(datalu, aes(x = Year)) +
    geom_line(aes(y = CA  , colour = "CA"), size = 1) +
    geom_line(aes(y = PR  , colour = "PR"), size = 1) +
    geom_line(aes(y = `NA`, colour = "NA"), size = 1) +
    scale_color_manual(values = scales::hue_pal()(3)) +
    theme_light() +
    labs(y = "Value",
         colour = "Series")

```{r}
ggplot(data, aes(x=time)) +
  geom_line(aes(y=temp), col="red") +
  geom_line(aes(y=rh), col="blue") +
  geom_line(aes(y=dewpt), col="black") +
  labs(x="Date") 

# change data format
datal <- data %>% 
  mutate(notice = ifelse(dewpt>rh, "warning", "clear")) %>% 
  pivot_longer(c(temp, rh, dewpt), names_to = 'measure', values_to = 'value') 

datal <- data %>% 
  mutate(notice = ifelse(dewpt>rh, 1, 0)) %>% 
  pivot_longer(c(temp, rh, dewpt), names_to = 'measure', values_to = 'value') 


# get start/end point of consecutive values for getting the shaded area
shade <- setDT(datal)[,id := rleid(notice)][
  ,.SD[.N > 1 & notice == 1],id][
  ,.(start=date[1],end=date[.N], run=.N),id]

# transform to POSIXct object for plotting
shade$startd <- as.POSIXct(shade$start)
shade$endd <- as.POSIXct(shade$end)

# plot all lines
# add shaded area where dew pt > RH 
ggplot() +
  geom_line(data=datal, aes(x=time, y=value, col=measure, group=measure), size=1) +
  theme_classic() +
  scale_color_manual(name="Measure", limits=c("temp", "rh", "dewpt"), labels = c("Temperature", "RH", "Dew point"), values = c("red", "blue", "black")) + 
  labs(x="Date", y="Value") +
   geom_rect(data=shade, 
            mapping=aes(xmin=startd, xmax=endd, ymin=0, ymax=100), fill='red', alpha=0.2)
```




##
Filtering plots 

Filter out by date range
```{r}
data %>% 
  filter(date > "2023-10-06") %>% 
  ggplot(aes(x=time, y=rh)) +
  geom_line()
```